\chapter{AI Methodology}
\label{ch:ai_methodology}

The core methodological contribution of this thesis is the design and deployment of an autonomous, multi-agent AI framework capable of end-to-end scientific validation. This framework transforms a traditional static code repository into a dynamic "validation workspace."

\section{System Architecture and Orchestration}
The framework is orchestrated using the DeerFlow platform, which coordinates multi-agent reasoning. Agents are equipped with specific toolchains via the Model Context Protocol (MCP), allowing them to interface with both local execution environments and remote scientific databases. This enables a seamless transition from literature retrieval to symbolic mathematics and numerical fitting.

\section{Retrieval-Augmented Generation for Scientific Validation}
A central component of our methodology is an Onyx-backed Retrieval-Augmented Generation (RAG) system. This system interfaces directly with academic APIs, including Semantic Scholar, arXiv, and Scite. By automating citation network analysis, the framework can evaluate the consensus around specific physical assumptions (e.g., the validity of the $\eta \to y$ Jacobian transformation at high $p_T$). This ensures that the baseline formulas used in our validation are historically and statistically grounded.

\section{Automated Symbolic and Numerical Pipeline}
The computational validation is split into two layers:
\begin{enumerate}
    \item \textbf{Symbolic Validation:} The framework parses draft manuscript equations and uses \texttt{SymPy} to symbolically verify Lorentz covariance, integration limits, and thermodynamic consistency (e.g., recovering the Cooper-Frye static limit as flow velocity $U \to 0$).
    \item \textbf{Numerical Fitting:} After symbolic verification, the equations are compiled into a numerical pipeline (\texttt{fitting\_pipeline.py}) utilizing the \texttt{iminuit} optimizer. The pipeline is designed to enforce rigorous statistical checks, including the computation of the covariance matrix, parameter correlations, residuals, and $\chi^2/ndf$. 
\end{enumerate}

This dual-layer approach prevents unphysical models (e.g., those suffering from intrinsic over-parameterization) from being accepted based on superficially reasonable visual fits.
